### 2018-05-22 SI.Rmd
genetic algorithm stuff

---
title: "Component response rate variation drives stability in large complex systems"
author: "Brad Duthie"
bibliography: references.bib
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
subtitle: Supporting information
biblio-style: apalike
---

```{r, echo = FALSE}
source(file = "R/sim_mat.R");
source(file = "R/plot_figs.R");
```

Contents
================================================================================

**Note: All code and data are publicly available on [GitHub](https://github.com/bradduthie/random_matrix_stability)**

- [Code and simulations underlying Fig. 1](#Fig1) 
- [Code and simulations underlying Fig. 2](#Fig2) 
- [Stability across increasing $S$](#Incr_S}) 
- [Stability of ecological networks](#ecological) 
    - [Competitor networks](#competition) 
    - [Mutualist networks](#mutualism) 
    - [Predator-prey networks](#pred-prey) 
- [Different connectance (C) values](#connectance)
    - [C = 0.3](#connect3)
    - [C = 0.5](#connect5)
    - [C = 0.7](#connect7)
    - [C = 0.9](#connect9)
- [Different distributions of $\gamma$](#gam_dist)
- [Feasilbility of complex systems](#Feasibility) 
- [Genetic algorithm](#ga)

Code and simulations underlying Fig. 1 {#Fig1}
================================================================================

The sample $M$ used for the eigenvalue distributions in Fig. 1 of the text is available on [GitHub](https://github.com/bradduthie/random_matrix_stability/tree/master/sim_results/bi_gamma), and was produced by running the following function.

```{r}
find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
    while(iters > 0){
        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
        A_mat  <- matrix(data = A_dat, nrow = S);
        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
        A_mat  <- A_mat * C_mat;
        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
        mu_gam <- mean(gammas);
        diag(A_mat) <- -1;
        A1     <- gammas * A_mat;
        A0     <- mu_gam * A_mat;
        A0_e   <- eigen(A0)$values;
        A0_r   <- Re(A0_e);
        A0_i   <- Im(A0_e);
        A1_e   <- eigen(A1)$values;
        A1_r   <- Re(A1_e);
        A1_i   <- Im(A1_e);
        if(max(A0_r) >= 0 & max(A1_r) < 0){
            return(list(A0 = A0, A1 = A1));
            break;
        }
        print(iters);
        iters <- iters - 1;
    }
}
```

The above function terminates when a matrix $M$ is found that is not stable when all component response rates are set to $\gamma = 1$, but is stable when half of component response rates are $1.95$ and half are $0.05$. The function is used to illustrate the concept of how fast versus slow component responses can cause a system to become stable. Simulations were run for `iter = 1000000`, but terminated once an acceptable `A0` and `A1` were found. The code below plots the eigenvalue distributions of `A0` and `A1` in panels **a** and **b**, respectively.

```{r, echo = FALSE}
A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
A0 <- as.matrix(A0[,-1]);
A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
A1 <- as.matrix(A1[,-1]);

A0_e   <- eigen(A0)$values;
A0_r   <- Re(A0_e);
A0_i   <- Im(A0_e);
A1_e   <- eigen(A1)$values;
A1_r   <- Re(A1_e);
A1_i   <- Im(A1_e);

A0_vm       <- A0;
diag(A0_vm) <- NA;
A0vec       <- as.vector(t(A0_vm));
A0vec       <- A0vec[is.na(A0vec) == FALSE];
A1_vm       <- A1;
diag(A1_vm) <- NA;
A1vec       <- as.vector(t(A1_vm));
A1vec       <- A1vec[is.na(A1vec) == FALSE];
fhalf       <- 1:(0.5*length(A1vec));
shalf       <- (0.5*length(A1vec)+1):length(A1vec);
```

The plot itself can be recreated with the code below.

```{r, fig.height = 6, fig.width = 9}
par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
vl <- seq(from = 0, to = 2*pi, by = 0.001);
A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
text(x = -3.5, y = 2.25, labels = "a", cex = 2);
points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "grey");
points(A0_r, A0_i, pch = 4, cex = 0.7);

plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
     col = "black", yaxt = "n");

vl <- seq(from = 0, to = 2*pi, by = 0.001);
A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");

points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7);

text(x = -3.5, y = 2.25, labels = "b", cex = 2);
mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
```

To find out how frequently $M$ was stable given that all $\gamma = 1$ versus $\gamma = \{1.95, 0.05\}$, the function below was created.

```{r}
stab_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
    ress     <- matrix(data = 0, nrow = iters, ncol = 2);
    A0_count <- 0;
    A1_count <- 0;
    while(iters > 0){
        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
        A_mat  <- matrix(data = A_dat, nrow = S);
        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
        A_mat  <- A_mat * C_mat;
        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
        mu_gam <- mean(gammas);
        diag(A_mat) <- -1;
        A1     <- gammas * A_mat;
        A0     <- mu_gam * A_mat;
        A0_e   <- eigen(A0)$values;
        A0_r   <- Re(A0_e);
        A0_i   <- Im(A0_e);
        A1_e   <- eigen(A1)$values;
        A1_r   <- Re(A1_e);
        A1_i   <- Im(A1_e);
        if(max(A0_r) < 0){
            ress[iters, 1] <- 1;
            A0_count       <- A0_count + 1;
            }
        if(max(A1_r) < 0){
            ress[iters, 2] <- 1;
            A1_count       <- A1_count + 1;
        }
        print(c(iters, A0_count, A1_count));
        iters <- iters - 1;
    }
    return(ress);
}
```

```{r, echo = FALSE}
pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
pr_st       <- pr_st[,-1];
```

The function above was run for `iters = 1000000`, and the resulting matrix `ress` was returned. Each row of `ress` represents a single $M$ given $\gamma = 1$ (column 1) versus $\gamma = \{1.95, 0.05\}$ (column 2). Values of 0 indicate that $M$ was found to be unstable (at least one real component of its eigenvalues greater than or equal to zero), whereas values of 1 indicate that $M$ was found to be stable (all real components of eigenvalues are negative). The frequencies of stable $M$ were `r sum(pr_st[,1])` given $\gamma = 1$ and `r sum(pr_st[,2])` given $\gamma = \{1.95, 0.05\}$, as reported in the main text and legend of Fig. 1 (raw data are [available on GitHub](https://github.com/bradduthie/random_matrix_stability/blob/master/sim_results/bi_gamma/bi_pr_st.csv)).

Code and simulations underlying Fig. 2 {#Fig2}
===============================================================================

Figure 2 of the main text shows eigenvalue distributions in a system where $S = 1000$, $C = 1$, and $\sigma = 0.4$. Eigenvalues can be reproduced using the code below for when $\gamma = 1$ (panel a) and $\gamma \sim \mathcal{U}(0, 2)$ (panel b).

```{r}
A_comp <- NULL;
A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
A_mat  <- matrix(data = A_dat, nrow = 1000);
C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
A_mat     <- A_mat * C_mat;
gammas <- runif(n = 1000, min = 0, max = 2);
mu_gam <- mean(gammas);
diag(A_mat) <- -1;
A1     <- gammas * A_mat;
A0     <- mu_gam * A_mat;
A0_e   <- eigen(A0)$values;
A0_r   <- Re(A0_e);
A0_i   <- Im(A0_e);
A1_e   <- eigen(A1)$values;
A1_r   <- Re(A1_e);
A1_i   <- Im(A1_e);

A0_vm       <- A0;
diag(A0_vm) <- NA;
A0vec       <- as.vector(A0_vm);
A0vec       <- A0vec[is.na(A0vec) == FALSE];
A1_vm       <- A1;
diag(A1_vm) <- NA;
A1vec       <- as.vector(A1_vm);
A1vec       <- A1vec[is.na(A1vec) == FALSE];
```

The code below reproduces the figure itself.

```{r, fig.height = 6, fig.width = 9}
par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
vl <- seq(from = 0, to = 2*pi, by = 0.001);
x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
text(x = -15.5, y = 19, labels = "a", cex = 2);
points(x = x0, y = y0, type = "l", lwd = 3);
points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
     yaxt = "n");
text(x = -15.5, y = 19, labels = "b", cex = 2);
points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
```

Stability across increasing $S$ {#Incr_S}
===============================================================================

```{r, echo = FALSE}
dat <- read.csv(file = "sim_results/C_1/random_all.csv");
dat <- dat[,-1];
```

The table below shows the results for all simulations of random $M$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$. In this table, the `A0` refers to matrices where $\gamma = 1$, while `A1` refers to matrices after $Var(\gamma)$ is added and $\gamma \sim \mathcal{U}(0, 2)$. Each row summarises data for a given $S$ over 1 million randomly simulated $M$ (`A0` and `A1`). The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 1 million). Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable. The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.

```{r, echo = FALSE}
library(knitr);
d_tab <- dat[,1:7];
kable(d_tab);
```

The results underlying this table were produced with the `rand_gen_var` function below.

```{r}
rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
    tot_res <- NULL;
    fea_res <- NULL;
    for(i in 2:max_sp){
        iter           <- iters;
        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
        while(iter > 0){
            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
            A0       <- species_interactions(mat = A0, type = int_type);
            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
            A0       <- A0 * C_mat;
            diag(A0) <- -1;
            gam1     <- runif(n = i, min = 0, max = 2);
            A1       <- A0 * gam1;
            A0       <- A0 * mean(gam1);
            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
            if(A0_stb == TRUE){
                tot_res[[i-1]][iter, 1] <- 1;
            }
            if(A1_stb == TRUE){
                tot_res[[i-1]][iter, 2] <- 1;
            }
            if(A0_fea == TRUE){
                fea_res[[i-1]][iter, 1] <- 1;
            }
            if(A1_fea == TRUE){
                fea_res[[i-1]][iter, 2] <- 1;
            }
            iter    <- iter - 1;
        }
        print(i);
    }
    all_res <- summarise_randmat(tot_res = tot_res, fea_res = fea_res);
    return(all_res);
}
```

The above function calls the two functions `species_interactions` and `summarise_randmat`, which are provided below.

```{r}
species_interactions <- function(mat, type = 0){
    if(type == 1){
        mat[mat > 0] <- -1*mat[mat > 0];
    }
    if(type == 2){
        mat[mat < 0] <- -1*mat[mat < 0];
    }
    if(type == 3){
        for(i in 1:dim(mat)[1]){
            for(j in 1:dim(mat)[2]){
                if(mat[i, j] * mat[j, i] > 0){
                    mat[j, i] <- -1 * mat[j, i];
                }
            }
        }
    }
    return(mat);
}

summarise_randmat <- function(tot_res, fea_res){
    sims    <- length(tot_res);
    all_res <- matrix(data = 0, nrow = sims, ncol = 13);
    for(i in 1:sims){
        all_res[i, 1]  <- i + 1;
        # Stable and unstable
        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
        # Stabilised and destabilised
        all_res[i, 6] <- sum(tot_res[[i]][,1] == FALSE & 
                                  tot_res[[i]][,2] == TRUE);
        all_res[i, 7] <- sum(tot_res[[i]][,1] == TRUE & 
                                  tot_res[[i]][,2] == FALSE);
        # Feasible and infeasible
        all_res[i, 8]  <- sum(fea_res[[i]][,1] == FALSE);
        all_res[i, 9]  <- sum(fea_res[[i]][,1] == TRUE);
        all_res[i, 10]  <- sum(fea_res[[i]][,2] == FALSE);
        all_res[i, 11]  <- sum(fea_res[[i]][,2] == TRUE);
        # Feased and defeased
        all_res[i, 12] <- sum(fea_res[[i]][,1] == FALSE & 
                                  fea_res[[i]][,2] == TRUE);
        all_res[i, 13] <- sum(fea_res[[i]][,1] == TRUE & 
                                  fea_res[[i]][,2] == FALSE);
    }
    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
                "A1_stabilised", "A1_destabilised", "A0_infeasible", 
                "A0_feasible", "A1_infeasible", "A1_feasible", 
                "A1_made_feasible", "A1_made_infeasible");
    colnames(all_res) <- cnames;
    return(all_res);
}
```

Note that feasibility results were ommited for the table above, but are [reported below](#Feasibility).

Stability of ecological networks {#ecological}
================================================================================

While the foundational work of May[@May1972] applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species. In these networks, the matrix $M$ is interpreted as a community matrix and each row and column is interpreted as a single species. The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $M_{ij}$, meaning that $M$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015a; @Tang2014b]. While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $M$ are negative), mutualist networks (all off-diagonal elements of $M$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2011; @Allesina2012; @Allesina2015a; @Tang2014b]. In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising.

I investigate competitor, mutualist, and predator-prey networks following Allesina et al.[@Allesina2012]. To create these networks, I first generated a random matrix $M$, then changed the elements of $M$ accordingly. If $M$ was a competitive network, then the sign of any positive off-diagonal elements was reversed to be negative. If $M$ was a mutualist network, then the sign of any positive off-diagonal elements was reversed to be positive. And if $M$ was a predator-prey network, then all $i$ and $j$ pairs of elements were checked; any pairs of the same sign were changed so that one was negative and the other was positive. The `species_interaction` function used to do this is below.

```{r}
species_interactions <- function(mat, type = 0){
    if(type == 1){
        mat[mat > 0] <- -1*mat[mat > 0];
    }
    if(type == 2){
        mat[mat < 0] <- -1*mat[mat < 0];
    }
    if(type == 3){
        for(i in 1:dim(mat)[1]){
            for(j in 1:dim(mat)[2]){
                if(mat[i, j] * mat[j, i] > 0){
                    mat[j, i] <- -1 * mat[j, i];
                }
            }
        }
    }
    return(mat);
} # Note: -1 values are added in the diagonal later
```

This function was applied to all created matrices $M$, then the number of stable $M$ matrices was estimated [exactly as it was](#Incr_S) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million. The following tables for restricted ecological communities can therefore be compared with the random $M$ [results above](#Incr_S). As with the [results above](#Incr_S), in the tables below, `A0` refers to matrices when $\gamma = 1$ and `A1` refers to matrices after $Var(\gamma)$ is added. The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 100000). Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable. The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.

<a name="competition">**Competition**</a>

Results for competitor interaction networks are shown below

```{r, echo = FALSE}
cdat   <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
cdat   <- cdat[,-1];
cd_tab <- cdat[,1:7];
kable(cd_tab);
```

<a name="mutualism">**Mutualism**</a>

Results for mutualist interaction networks are shown below

```{r, echo = FALSE}
mdat   <- read.csv(file = "sim_results/ecology/mutualism_C_1.csv");
mdat   <- mdat[,-1];
md_tab <- mdat[,1:7];
kable(md_tab);
```

<a name="pred-prey">**Predator-prey**</a>

Results for predator-prey interaction networks are shown below

```{r, echo = FALSE}
pdat   <- read.csv(file = "sim_results/ecology/pred-prey_C_1.csv");
pdat   <- pdat[,-1];
pd_tab <- pdat[,1:7];
kable(pd_tab);
```

Overall, as expected[@Allesina2012], predator-prey communities are relatively stable while mutualist communties are highly unstable. But interestingly, while $Var(\gamma)$ stabilises predator-prey and competitor communities, it does not stabilise mutualist communities. This is unsurprising because purely mutualist communities are characterised by a very positive[@Allesina2012] leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.

Different connectance (C) values {#connectance}
================================================================================

In the main text, for simplicity, I assumed connectance values of $C = 1$, meaning that all off-diagonal elements of a matrix $M$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2})$ where $\sigma = 0.4$. Here I present four tables showing the number of stable communities given $C = \{0.3, 0. 5, 0.7, 0.9 \}$. In all cases, uniform variation in component response time ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma = 1$). In contrast to the main text, 100000 rather than 1 million $M$ were simulated. As with the results on [stability with increasing $S$](#Incr_S) shown above, in the tables below `A0` refers to matrices when $\gamma = 1$, and `A1` refers to matrices after $Var(\gamma)$ is added. The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 100000). Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable. The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.

<a name="connect3">**Connectance $C = 0.3$**</a>

```{r, echo = FALSE}
C3dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt3.csv");
C3dat   <- C3dat[,-1];
C3d_tab <- C3dat[,1:7];
kable(C3d_tab);
```

<a name="connect5">**Connectance $C = 0.5$**</a>

```{r, echo = FALSE}
C5dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt5.csv");
C5dat   <- C5dat[,-1];
C5d_tab <- C5dat[,1:7];
kable(C5d_tab);
```

<a name="connect7">**Connectance $C = 0.7$**</a>

```{r, echo = FALSE}
C7dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt7.csv");
C7dat   <- C7dat[,-1];
C7d_tab <- C7dat[,1:7];
kable(C7d_tab);
```

<a name="connect9">**Connectance $C = 0.9$**</a>

```{r, echo = FALSE}
C9dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt9.csv");
C9dat   <- C9dat[,-1];
C9d_tab <- C9dat[,1:7];
kable(C9d_tab);
```

Different distributions of $\gamma$ {#gam_dist}
================================================================================

In the main text, I considered a uniform distribution of component response rates $\gamma \sim \mathcal{U}(0, 2)$. The number of unstable and stable $M$ matrices are reported in [a table above](#Incr_S) across different values of $S$. Here I show complementary results for three different distributions including an exponential, beta, and gamma distribution of $\gamma$ values. The shape of these distributions is shown in the figure below.

********************************************************************************

**Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.** The stabilities of simulated complex systems with these $\gamma$ distributions are compared to otherwise identical complex systems with a fixed component response rate of $\gamma = 1$ across different system sizes ($S$; i.e., component numbers) given a unit $\gamma$ standard deviation ($\sigma_{\gamma} = 1$) for b-d. Distributions are as follows: (a) uniform, (b) exponential, (c) beta ($\alpha = 0.5$ and $\beta = 0.5$), and (d) gamma ($k = 2$ and $\theta = 2$). Each panel shows 1 million randomly generated $\gamma$ values.

```{r, eval = TRUE, echo = FALSE, fig.height = 7, fig.width = 6}
make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
    if(distribution == 0){
        dat          <- rep(x = mn, times = nn);
    }
    if(distribution == 1){
        mval         <- 2;
        dat          <- runif(n = nn, min = 0, max = mval);
    }
    if(distribution == 2){
        dat          <- rexp(n = nn);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat - min(dat) + min(abs(dat));
    }
    if(distribution == 3){
        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat - min(dat) + min(abs(dat));
    }
    if(distribution == 4){
        dat          <- rgamma(n = nn, shape = 2, scale = 2);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat - min(dat) + min(abs(dat));
    }
    return(dat);
}

show_gammas <- function(nn = 1000000, sdd = 1, mn = 10){
    y1 <- make_gammas(nn = nn, distribution = 1, sdd = sdd, mn = mn);
    y2 <- make_gammas(nn = nn, distribution = 2, sdd = sdd, mn = mn);
    y3 <- make_gammas(nn = nn, distribution = 3, sdd = sdd, mn = mn);
    y4 <- make_gammas(nn = nn, distribution = 4, sdd = sdd, mn = mn);
    par(mfrow = c(2, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
    h1 <- hist(y1, breaks = 1000, plot = FALSE);
    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
         ylim = c(0, max(h1$counts)*1.2));
    mtext("a", adj = 0, line = -2.3, cex = 2, 
          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
    box();
    h2 <- hist(y2, breaks = 1000, plot = FALSE);
    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
         ylim = c(0, max(h2$counts)*1.2));
    mtext("b", adj = 0, line = -2.3, cex = 2, 
          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
    box();
    h3 <- hist(y3, breaks = 1000, plot = FALSE);
    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
         ylim = c(0, max(h3$counts)*1.2));
    mtext("c", adj = 0, line = -2.3, cex = 2, 
          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
    box();
    h4 <- hist(y4, breaks = 1000, plot = FALSE);
    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
         ylim = c(0, max(h4$counts)*1.2));
    mtext("d", adj = 0, line = -2.3, cex = 2, 
          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
    box();
    mtext(expression(paste("Component ",gamma," value")),
          outer=TRUE,side=1,line=3.0,cex=2);
    mtext(expression(paste("Relative frequency")),
          outer=TRUE,side=2,line=2.5,cex=2);
}

show_gammas();
```

********************************************************************************

The same 100000 $M$ matrices were used to investigate stability when applying each of these different distributions of $\gamma$ values. The table below shows the number of $M$ that were unstable (_unst) and stable (_stbl) for the exponential (Exp), beta, and gamma distributions.

```{r, echo = TRUE}
fordist   <- read.csv(file = "sim_results/different_distr/four_distr_rand.csv");
kable(fordist);
```

In comparison to the uniform distribution (a), proportionally fewer random systems are found with the exponential distribution (b), while more are found with the beta (c) and gamma (d) distributions.

Genetic algorithm {#ga}
================================================================================

Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluate for each unique $M$. This is technically impossible because any $\gamma_{i}$ can take any real value between $0-2$, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore. Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013]. In this case, the practical solution is finding vectors of $\gamma$ that decrease the most positive real eigenvalue of $M$. The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below 20 generations pass or the fitness increase from one generation to the next is below a certain criteria). The genetic algorithm relies on five functions. The first outer function `Evo_rand_gen_var` runs all of the simulations (`from` and `to` refer to $S$ values, and `iters` refers to the number of $M$ to try for each $S$).

```{r}
Evo_rand_gen_var <- function(from, to, iters, int_type = 0, rmx = 0.4, C = 1){
    tot_res <- NULL;
    fea_res <- NULL;
    if(from >= to){
        stop("Argument 'from' must be less than argument 'to'");
    }
    for(i in from:to){
        nn             <- i;
        A1_stt         <- 0;
        A2_stt         <- 0;
        A1_fet         <- 0;
        A2_fet         <- 0;
        iter           <- iters;
        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 3);
        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
        while(iter > 0){
            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
            A0       <- species_interactions(mat = A0, type = int_type);
            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
            A0       <- A0 * C_mat;
            diag(A0) <- -1;
            gam1     <- runif(n = i, min = 0, max = 2);
            A1       <- A0 * gam1;
            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
            A1_stb   <- rand_mat_ga(A1);
            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
            if(A0_stb == TRUE){
                tot_res[[i-1]][iter, 1] <- 1;
            }
            if(A1_stb == TRUE){
                tot_res[[i-1]][iter, 2] <- 1;
            }
            if(A0_fea == TRUE){
                fea_res[[i-1]][iter, 1] <- 1;
            }
            if(A1_fea == TRUE){
                fea_res[[i-1]][iter, 2] <- 1;
            }
            iter    <- iter - 1;
        }
        print(i);
    }
    all_res <- summarise_randmat(tot_res = tot_res, fea_res = fea_res);
    return(all_res);
}
```

Note that `Evo_rand_gen_var` calls three custom sub-functions, `species_interactions`, `rand_mat_ga`, and `summarise_randmat`. The first simply allows for non-random interactions between components (e.g., modelling [ecological interactions](#ecological) of random, competition, mutualism, or predator-prey).

```{r}
species_interactions <- function(mat, type = 0){
    if(type == 1){
        mat[mat > 0] <- -1*mat[mat > 0];
    }
    if(type == 2){
        mat[mat < 0] <- -1*mat[mat < 0];
    }
    if(type == 3){
        for(i in 1:dim(mat)[1]){
            for(j in 1:dim(mat)[2]){
                if(mat[i, j] * mat[j, i] > 0){
                    mat[j, i] <- -1 * mat[j, i];
                }
            }
        }
    }
    return(mat);
}
```

The sub-function `rand_mat_ga` does the work of the genetic algorithm, searching for $\gamma$ vectors that are stabilising.

```{r}
rand_mat_ga <- function(A1, max_it = 20, converg = 0.01){
    nn       <- dim(A1)[1];
    rind     <- runif(n = nn*1000, min = 0, max = 1);
    inds     <- matrix(data = rind, nrow = 1000, ncol = nn);
    lastf    <- -10;
    ccrit    <- 10;
    find_st  <- 0;
    iter     <- max_it;
    while(iter > 0 & find_st < 1 & ccrit > converg){
        ivar  <- rep(x = 0, length = dim(inds)[1]);
        ifit  <- rep(x = 0, length = dim(inds)[1]);
        isst  <- rep(x = 0, length = dim(inds)[1]);
        for(i in 1:dim(inds)[1]){
            ifit[i] <- -1*max(Re(eigen(inds[i,]*A1)$values));
            ivar[i] <- var(inds[i,]);
            isst[i] <- max(Re(eigen(inds[i,]*A1)$values)) < 0;
        }
        most_fit <- order(ifit, decreasing = TRUE)[1:20];
        parents  <- inds[most_fit,];
        new_gen  <- matrix(data = t(parents), nrow = 1000, ncol = nn, 
                           byrow = TRUE);
        mu_dat   <- rbinom(n = nn*1000, size = 1, prob = 0.2);
        mu_dat2  <- rnorm(n = nn*1000, mean = 0, sd = 0.02);
        mu_dat2[mu_dat2 < 0] <- -mu_dat2[mu_dat2 < 0];
        mu_dat2[mu_dat2 > 2] <- 2;
        mu_dat3  <- mu_dat * mu_dat2;
        mu_mat   <- matrix(data = mu_dat3, nrow = 1000, ncol = nn);
        new_gen  <- new_gen + mu_mat;
        new_gen  <- crossover(inds = new_gen, pr = 0.1);
        inds     <- new_gen;
        find_st  <- max(isst);
        newf     <- mean(ifit);
        ccrit    <- newf - lastf;
        lastf    <- newf;
        iter     <- iter - 1;
    }
    if(find_st == 1){
        s_row <- which(isst == 1)[1];
        writt <- c(nn, inds[s_row,]);
        cat(writt, file = "evo_out.txt", append = TRUE);
        cat("\n", file = "evo_out.txt", append = TRUE);
    }
    return(find_st);
}
```

The while loop continues until either `iter` generations have occured, a solution $\gamma$ vector is found that results in all $\Re(\lambda) < 0$, or some criteria of minimum fitness increase is observed. Within the genetic algorithm, $\gamma$ values are mutated, crossover occurs between $\gamma$ vectors, and there is selection occurs in each generation such that the 20 $\gamma$ vectors that produce the lowest maximum $\Re(\lambda)$ are allowed to have 50 offspring each. In mutation, any values that mutate below zero are multiplied by $-1$, and any values that mutate above 2 are set to 2. Note also that if a solution is found, then one such $\gamma$ vector causing stability is printed to a file.

Crossover occurs in the `crossover` function below.

```{r} 
crossover <- function(inds, pr = 0.1){
    crossed <- floor(dim(inds)[1] * pr);
    cross1  <- sample(x = 1:dim(inds)[1], size = crossed);
    cross2  <- sample(x = 1:dim(inds)[1], size = crossed);
    for(i in 1:length(cross1)){
        fromv   <- sample(x = 1:dim(inds)[2], size = 1);
        tov     <- sample(x = 1:dim(inds)[2], size = 1);
        temp                   <- inds[cross1[i],fromv:tov];
        inds[cross1[i],fromv:tov] <- inds[cross2[i],fromv:tov];
        inds[cross2[i],fromv:tov] <- temp;
    }
    return(inds);
}
```

After all $M$ are simulated in `Evo_rand_gen_var`, the `summarise_randmat` formats the data into a table.

```{r}
summarise_randmat <- function(tot_res, fea_res){
    sims    <- length(tot_res);
    all_res <- matrix(data = 0, nrow = sims, ncol = 10);
    for(i in 1:sims){
        unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
        stables   <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == TRUE;
        unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
        stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
        non_feas  <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == FALSE;
        feasibl   <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == TRUE;
        unfeased  <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == FALSE;
        feased    <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == TRUE;
        foundd    <- tot_res[[i]][,3] == TRUE;
        all_res[i, 1]  <- i + 1;
        all_res[i, 2]  <- sum(unstables);
        all_res[i, 3]  <- sum(stables);
        all_res[i, 4]  <- sum(unstabled);
        all_res[i, 5]  <- sum(stabled);
        all_res[i, 6]  <- sum(non_feas);
        all_res[i, 7]  <- sum(feasibl);
        all_res[i, 8]  <- sum(unfeased);
        all_res[i, 9]  <- sum(feased);
        all_res[i, 10] <- sum(foundd);
    }
    return(all_res);
}
```

Stability results from this table are shown below.

```{r, echo = FALSE}
evo_res   <- read.csv(file = "sim_results/evolved/evo_results.csv");
kable(evo_res);
```

The distribution of one of the $\gamma$ vectors at $S = 39$ is shown below (all are available on [GitHub]()).

**Work in progress**

Feasilbility of complex systems {#Feasibility}
================================================================================

**Work in progress**

It also would be useful to look at feasibility criteria established by @Dougoud2018, who very recently made the point that some of May and Allesina's criteria allows for negative species densities when stable. Feasibility criteria are as follows,

$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
The above is not nearly as nasty as it looks, especially because it is entirely reasonable to simply use convenient values to parameterise it. The variable $x^{*}$ is just the vector of species abundances at equilibrium (we need all of them to be positive). The matrix $I$ is just the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is just the strength of intraspecific competition -- we can just set this to $\theta = -1$ (as others have). The variable $C$ is just the connectance of the community, which we will also set to 1 for convenience (lower values would mean that some species don't interact with one another, corresponding to an off-diagonal matrix element of 0). And $\delta$ just affects the strength of interactions -- which we can set to $\delta = 0$ for strong interactions. Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off-diagonal effecting species interactions,

$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$

To check the feasibility criteria, all that needs to be done is to invert $(\theta I + A)$ and multiply the matrix by the vector of growth rates $r$, which we can also just set to 1. So really, we're just multiplying $-(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive. This isn't much extra work, but it will probably go a long way toward satisfying any reviewers familiar with @Dougoud2018.

References
================================================================================