### 2018-06-08 notebook/SI.tex
Fix Issue #2 in the SI; dummy index removed from all ecological tables

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{setspace}
\setstretch{1}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Component response rate variation drives stability in large complex systems},
            pdfauthor={A. Bradley Duthie},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Component response rate variation drives stability in large complex
systems}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Supplemental Information}
  \author{A. Bradley Duthie}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{Biological and Environmental Sciences, University of Stirling, Stirling,
UK, FK9 4LA
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\linenumbers
\bibliographystyle{amnatnat}

\begin{document}
\maketitle

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{This supplemental information supports the manuscript
``Component response rate variation drives stability in large complex
systems'' with all of the code required to recreate the analysis in the
main text, and with additional analyses to support its conclusions. All
text, code, and data underlying this manuscript are publicly available
on \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
part of the RandomMatrixStability package.}

The
\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability
package} includes all functions and tools for recreating the text, this
supplemental information, and running all code; additional documentation
is also provided for functions as part of the package. The
RandomMatrixStability package package is available on
\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
download it, the
\href{https://cran.r-project.org/web/packages/devtools/index.html}{\texttt{devtools}
library} is needed.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{);}
\KeywordTok{library}\NormalTok{(devtools);}
\end{Highlighting}
\end{Shaded}

The code below installs the RandomMatrixStability package using
devtools.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/RandomMatrixStability"}\NormalTok{);}
\end{Highlighting}
\end{Shaded}

While downloading this package is recommended, all relevant code is also
reproduced below with explanation, so it is possible to recreate all
analyses using only this supplemental information.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\section{Supplemental Information table of
contents}\label{supplemental-information-table-of-contents}

\begin{itemize}
\tightlist
\item
  \protect\hyperlink{Fig1}{Code and simulations underlying Fig. 1}
\item
  \protect\hyperlink{Fig2}{Code and simulations underlying Fig. 2}
\item
  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
\item
  \protect\hyperlink{ecological}{Stability of ecological networks}

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{competition}{Competitor networks}
  \item
    \protect\hyperlink{mutualism}{Mutualist networks}
  \item
    \protect\hyperlink{pred-prey}{Predator-prey networks}
  \end{itemize}
\item
  \protect\hyperlink{connectance}{Different connectance (C) values}

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{connect3}{C = 0.3}
  \item
    \protect\hyperlink{connect5}{C = 0.5}
  \item
    \protect\hyperlink{connect7}{C = 0.7}
  \item
    \protect\hyperlink{connect9}{C = 0.9}
  \end{itemize}
\item
  \protect\hyperlink{gam_dist}{Different distributions of \(\gamma\)}
\item
  \protect\hyperlink{ga}{Genetic algorithm}
\item
  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
\end{itemize}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{Fig1}{\section{Code and simulations underlying Fig.
1}\label{Fig1}}

The sample \(M\) used for the eigenvalue distributions in Fig. 1 of the
text is available on
\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma}{GitHub},
and was produced by running the following function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{find_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) >=}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \KeywordTok{return}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1));}
            \NormalTok{break;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(iters);}
        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{\}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The above \texttt{find\_bgamma} function terminates when a matrix \(M\)
is found that is not stable when all component response rates are set to
\(\gamma = 1\), but is stable when half of component response rates are
\(1.95\) and half are \(0.05\). The function is used to illustrate the
concept of how fast versus slow component responses can cause a system
to become stable. Simulations were run for \texttt{iter\ =\ 1000000},
but terminated once an acceptable \texttt{A0} and \texttt{A1} were
found. The code below plots the eigenvalue distributions of \texttt{A0}
and \texttt{A1} in panels \textbf{a} and \textbf{b}, respectively. The
plot itself can be recreated with the function and code below.

\includegraphics{SI_files/figure-latex/unnamed-chunk-5-1.pdf}

To find out how frequently \(M\) was stable given that all
\(\gamma = 1\) versus \(\gamma = \{1.95, 0.05\}\), the function below
was created.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{stab_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
    \NormalTok{ress     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iters, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
    \NormalTok{A0_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
    \NormalTok{A1_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{ress[iters, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{A0_count       <-}\StringTok{ }\NormalTok{A0_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{ress[iters, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{A1_count       <-}\StringTok{ }\NormalTok{A1_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(}\KeywordTok{c}\NormalTok{(iters, A0_count, A1_count));}
        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(ress);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The above functions produced the \texttt{bi\_pr\_st} data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bi_pr_st <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"sim_results/bi_gamma/bi_pr_st.csv"}\NormalTok{);}
\NormalTok{pr_st    <-}\StringTok{ }\NormalTok{bi_pr_st[,-}\DecValTok{1}\NormalTok{];}
\end{Highlighting}
\end{Shaded}

The function \texttt{stab\_bgamma} was run for
\texttt{iters\ =\ 1000000}, and the resulting matrix \texttt{ress} was
returned. Each row of \texttt{ress} represents a single \(M\) given
\(\gamma = 1\) (column 1) versus \(\gamma = \{1.95, 0.05\}\) (column 2).
Values of 0 indicate that \(M\) was found to be unstable (at least one
real component of its eigenvalues greater than or equal to zero),
whereas values of 1 indicate that \(M\) was found to be stable (all real
components of eigenvalues are negative). The frequencies of stable \(M\)
were 1 given \(\gamma = 1\) and 32 given \(\gamma = \{1.95, 0.05\}\), as
reported in the main text and legend of Fig. 1 (raw data are
\href{https://github.com/bradduthie/RandomMatrixStability/blob/master/sim_results/bi_gamma/bi_pr_st.csv}{available
on GitHub}).

\hypertarget{Fig2}{\section{Code and simulations underlying Fig.
2}\label{Fig2}}

Figure 2 of the main text shows eigenvalue distributions in a system
where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). Eigenvalues can be
reproduced using the code below for when \(\gamma = 1\) (panel a) and
\(\gamma \sim \mathcal{U}(0, 2)\) (panel b). The function below
reproduces the figure.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot_Fig_2 <-}\StringTok{ }\NormalTok{function()\{}
    \NormalTok{A_comp <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
    \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{);}
    \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000} \NormalTok{*}\StringTok{ }\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \DecValTok{1}\NormalTok{);}
    \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{1000}\NormalTok{);}
    \NormalTok{A_mat     <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
    \NormalTok{gammas <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
    \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
    \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
    \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
    \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
    \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
    \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
    \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
    \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
    \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
    \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A0_vm);}
    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A1_vm);}
    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }
         \DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }
         \DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
    \NormalTok{x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
    \NormalTok{y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
    \NormalTok{x1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1));}
    \NormalTok{y1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{);}
    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{,}
         \DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{)}
    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
\NormalTok{\}}
\KeywordTok{plot_Fig_2}\NormalTok{();}
\end{Highlighting}
\end{Shaded}

\includegraphics{SI_files/figure-latex/unnamed-chunk-8-1.pdf}

\hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
\(S\)}{Stability across increasing S}}\label{IncrS}}

Figure 3 of the main text reports the number of stable random complex
systems found over 1 million iterations. The data used to make this
figure are read into R below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_1/random_all.csv"}\NormalTok{);}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[,-}\DecValTok{1}\NormalTok{]; }\CommentTok{# Extra row-indicating column removed}
\end{Highlighting}
\end{Shaded}

The table below shows the results for all simulations of random \(M\)
matrices at \(\sigma = 0.4\) and \(C = 1\) given a range of
\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
Each row summarises data for a given \(S\) over 1 million randomly
simulated \(M\) (\texttt{A0} and \texttt{A1}). The column
\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
unstable, and the column \texttt{A0\_stable} shows the number of
\texttt{A0} matrices that are stable (these two columns sum to 1
million). Similarly, the column \texttt{A1\_unstable} shows the number
of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
the number that are stable. The columns \texttt{A1\_stabilised} and
\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
stabilised or destabilised, respectively, by \(Var(\gamma)\).

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 293 & 999707 & 293 & 999707 & 0 & 0\tabularnewline
3 & 3602 & 996398 & 3609 & 996391 & 0 & 7\tabularnewline
4 & 14937 & 985063 & 15008 & 984992 & 0 & 71\tabularnewline
5 & 39289 & 960711 & 39783 & 960217 & 36 & 530\tabularnewline
6 & 78845 & 921155 & 80207 & 919793 & 389 & 1751\tabularnewline
7 & 133764 & 866236 & 136904 & 863096 & 1679 & 4819\tabularnewline
8 & 204112 & 795888 & 208241 & 791759 & 5391 & 9520\tabularnewline
9 & 288041 & 711959 & 291775 & 708225 & 12619 & 16353\tabularnewline
10 & 384024 & 615976 & 384931 & 615069 & 23153 & 24060\tabularnewline
11 & 485975 & 514025 & 481019 & 518981 & 35681 & 30725\tabularnewline
12 & 590453 & 409547 & 577439 & 422561 & 48302 & 35288\tabularnewline
13 & 689643 & 310357 & 669440 & 330560 & 57194 & 36991\tabularnewline
14 & 777496 & 222504 & 751433 & 248567 & 60959 & 34896\tabularnewline
15 & 850159 & 149841 & 821613 & 178387 & 58567 & 30021\tabularnewline
16 & 905057 & 94943 & 877481 & 122519 & 51255 & 23679\tabularnewline
17 & 943192 & 56808 & 919536 & 80464 & 40854 & 17198\tabularnewline
18 & 969018 & 30982 & 949944 & 50056 & 30102 & 11028\tabularnewline
19 & 984301 & 15699 & 970703 & 29297 & 20065 & 6467\tabularnewline
20 & 992601 & 7399 & 983507 & 16493 & 12587 & 3493\tabularnewline
21 & 996765 & 3235 & 991532 & 8468 & 7030 & 1797\tabularnewline
22 & 998693 & 1307 & 995567 & 4433 & 3884 & 758\tabularnewline
23 & 999503 & 497 & 997941 & 2059 & 1883 & 321\tabularnewline
24 & 999861 & 139 & 999059 & 941 & 899 & 97\tabularnewline
25 & 999964 & 36 & 999617 & 383 & 380 & 33\tabularnewline
26 & 999993 & 7 & 999878 & 122 & 121 & 6\tabularnewline
27 & 999995 & 5 & 999946 & 54 & 53 & 4\tabularnewline
28 & 1000000 & 0 & 999975 & 25 & 25 & 0\tabularnewline
29 & 1000000 & 0 & 999997 & 3 & 3 & 0\tabularnewline
30 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
31 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
matrices found is greater than 1 (compare column 5 to column 3), and
this ratio increases with increasing \(S\) (column 1). Hence, more
randomly created complex systems (\(M\)) are generated given variation
in \(\gamma\) than when \(\gamma = 1\). The results underlying this
table were produced with the \texttt{rand\_gen\_var} function below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\KeywordTok{mean}\NormalTok{(gam1);}
            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(i);}
    \NormalTok{\}}
    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
    \KeywordTok{return}\NormalTok{(all_res);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The above function calls the two functions
\texttt{species\_interactions} and \texttt{summarise\_randmat}, which
are provided below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
                \NormalTok{\}}
            \NormalTok{\}}
        \NormalTok{\}}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(mat);}
\NormalTok{\}}

\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{13}\NormalTok{);}
    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \CommentTok{# Stable and unstable}
        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \CommentTok{# Stabilised and destabilised}
        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \CommentTok{# Feasible and infeasible}
        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{11}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \CommentTok{# Feased and defeased}
        \NormalTok{all_res[i, }\DecValTok{12}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{13}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
    \NormalTok{\}}
    \NormalTok{cnames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{, }\StringTok{"A1_stable"}\NormalTok{, }
                \StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{, }\StringTok{"A0_infeasible"}\NormalTok{, }
                \StringTok{"A0_feasible"}\NormalTok{, }\StringTok{"A1_infeasible"}\NormalTok{, }\StringTok{"A1_feasible"}\NormalTok{, }
                \StringTok{"A1_made_feasible"}\NormalTok{, }\StringTok{"A1_made_infeasible"}\NormalTok{);}
    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\NormalTok{cnames;}
    \KeywordTok{return}\NormalTok{(all_res);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Note that feasibility results were ommited for the table above, but are
\protect\hyperlink{Feasibility}{reported below}.

\hypertarget{ecological}{\section{Stability of ecological
networks}\label{ecological}}

While the foundational work of
May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
to complex networks, much attention has been given specifically to
ecological networks of interacting species. In these networks, the
matrix \(M\) is interpreted as a community matrix and each row and
column is interpreted as a single species. The effect that the density
of any species \(i\) has on the population dynamics of species \(j\) is
found in \(M_{ij}\), meaning that \(M\) holds the effects of pair-wise
interactions between \(S\)
species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Tang2014b}{4}}.
While May's original
work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
randomly assembled communities, recent work has specifically looked at
more restricted ecological communities including competitive networks
(all off-diagonal elements of \(M\) are negative), mutualist networks
(all off-diagonal elements of \(M\) are positive), and predator-prey
networks (for any pair of \(i\) and \(j\), the effect of \(i\) on \(j\)
is negative and \(j\) on \(i\) is positive, or vice
versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Allesina2011}{5}}.
In general, competitor and mutualist networks tend to be unstable, while
predator-prey networks tend to be highly stabilising.

I investigated competitor, mutualist, and predator-prey networks
following Allesina et
al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. To create
these networks, I first generated a random matrix \(M\), then changed
the elements of \(M\) accordingly. If \(M\) was a competitive network,
then the sign of any positive off-diagonal elements was reversed to be
negative. If \(M\) was a mutualist network, then the sign of any
positive off-diagonal elements was reversed to be positive. And if \(M\)
was a predator-prey network, then all \(i\) and \(j\) pairs of elements
were checked; any pairs of the same sign were changed so that one was
negative and the other was positive. The \texttt{species\_interaction}
function used to do this is below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
                \NormalTok{\}}
            \NormalTok{\}}
        \NormalTok{\}}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(mat);}
\NormalTok{\} }\CommentTok{# Note: -1 values are added in the diagonal later}
\end{Highlighting}
\end{Shaded}

This function was applied to all created matrices \(M\), then the number
of stable \(M\) matrices was estimated \protect\hyperlink{IncrS}{exactly
as it was} in the main text for random matrices for values of \(S\) from
2 to 50 (100 in the case of the relatively more stable predator-prey
interactions), except that only 100000 random \(M\) were generated
instead of 1 million. This produced the data set below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/competition_C_1.csv"}\NormalTok{);}
\NormalTok{mdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/mutualism_C_1.csv"}\NormalTok{);}
\NormalTok{pdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/pred-prey_C_1.csv"}\NormalTok{);}
\end{Highlighting}
\end{Shaded}

The following tables for restricted ecological communities can therefore
be compared with the random \(M\) \protect\hyperlink{IncrS}{results
above} (but note that counts from systems with comparable probabilities
of stability will be an order of magnitude lower in the tables below due
to the smaller number of \(M\) matrices generated). As with the
\protect\hyperlink{IncrS}{results above}, in the tables below,
\texttt{A0} refers to matrices when \(\gamma = 1\) and \texttt{A1}
refers to matrices after \(Var(\gamma)\) is added. The column
\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
unstable, and the column \texttt{A0\_stable} shows the number of
\texttt{A0} matrices that are stable (these two columns sum to 100000).
Similarly, the column \texttt{A1\_unstable} shows the number of
\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
number that are stable. The columns \texttt{A1\_stabilised} and
\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
stabilised or destabilised, respectively, by \(Var(\gamma)\).

\textbf{Competition}

Results for competitor interaction networks are shown below

\begin{longtable}[]{@{}llllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised\tabularnewline
\midrule
\endhead
2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Mutualism}

Results for mutualist interaction networks are shown below

\begin{longtable}[]{@{}llllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised\tabularnewline
\midrule
\endhead
2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Predator-prey}

Results for predator-prey interaction networks are shown below

\begin{longtable}[]{@{}rrrrrr@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised\tabularnewline
\midrule
\endhead
2 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
3 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
4 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
5 & 1 & 99999 & 1 & 99999 & 0\tabularnewline
6 & 4 & 99996 & 4 & 99996 & 0\tabularnewline
7 & 2 & 99998 & 2 & 99998 & 0\tabularnewline
8 & 5 & 99995 & 5 & 99995 & 0\tabularnewline
9 & 20 & 99980 & 21 & 99979 & 0\tabularnewline
10 & 20 & 99980 & 22 & 99978 & 0\tabularnewline
11 & 38 & 99962 & 39 & 99961 & 0\tabularnewline
12 & 64 & 99936 & 66 & 99934 & 0\tabularnewline
13 & 87 & 99913 & 91 & 99909 & 0\tabularnewline
14 & 157 & 99843 & 159 & 99841 & 0\tabularnewline
15 & 215 & 99785 & 227 & 99773 & 0\tabularnewline
16 & 293 & 99707 & 310 & 99690 & 0\tabularnewline
17 & 383 & 99617 & 408 & 99592 & 0\tabularnewline
18 & 443 & 99557 & 473 & 99527 & 3\tabularnewline
19 & 642 & 99358 & 675 & 99325 & 4\tabularnewline
20 & 836 & 99164 & 887 & 99113 & 7\tabularnewline
21 & 1006 & 98994 & 1058 & 98942 & 10\tabularnewline
22 & 1153 & 98847 & 1228 & 98772 & 20\tabularnewline
23 & 1501 & 98499 & 1593 & 98407 & 30\tabularnewline
24 & 1841 & 98159 & 1996 & 98004 & 40\tabularnewline
25 & 2146 & 97854 & 2316 & 97684 & 58\tabularnewline
26 & 2643 & 97357 & 2809 & 97191 & 119\tabularnewline
27 & 3034 & 96966 & 3258 & 96742 & 158\tabularnewline
28 & 3690 & 96310 & 3928 & 96072 & 201\tabularnewline
29 & 4257 & 95743 & 4532 & 95468 & 290\tabularnewline
30 & 4964 & 95036 & 5221 & 94779 & 424\tabularnewline
31 & 5627 & 94373 & 5978 & 94022 & 452\tabularnewline
32 & 6543 & 93457 & 6891 & 93109 & 666\tabularnewline
33 & 7425 & 92575 & 7777 & 92223 & 818\tabularnewline
34 & 8540 & 91460 & 8841 & 91159 & 1071\tabularnewline
35 & 9526 & 90474 & 9842 & 90158 & 1337\tabularnewline
36 & 10617 & 89383 & 10891 & 89109 & 1624\tabularnewline
37 & 12344 & 87656 & 12508 & 87492 & 2021\tabularnewline
38 & 13675 & 86325 & 13877 & 86123 & 2442\tabularnewline
39 & 15264 & 84736 & 15349 & 84651 & 2870\tabularnewline
40 & 17026 & 82974 & 17053 & 82947 & 3363\tabularnewline
41 & 18768 & 81232 & 18614 & 81386 & 3905\tabularnewline
42 & 20791 & 79209 & 20470 & 79530 & 4579\tabularnewline
43 & 23150 & 76850 & 22754 & 77246 & 5217\tabularnewline
44 & 25449 & 74551 & 24184 & 75816 & 6285\tabularnewline
45 & 27702 & 72298 & 26464 & 73536 & 6754\tabularnewline
46 & 30525 & 69475 & 28966 & 71034 & 7646\tabularnewline
47 & 32832 & 67168 & 31125 & 68875 & 8487\tabularnewline
48 & 36152 & 63848 & 33865 & 66135 & 9479\tabularnewline
49 & 38714 & 61286 & 36242 & 63758 & 10125\tabularnewline
50 & 41628 & 58372 & 38508 & 61492 & 11036\tabularnewline
51 & 44483 & 55517 & 41023 & 58977 & 11704\tabularnewline
52 & 48134 & 51866 & 44287 & 55713 & 12573\tabularnewline
53 & 51138 & 48862 & 46721 & 53279 & 13223\tabularnewline
54 & 54261 & 45739 & 49559 & 50441 & 13757\tabularnewline
55 & 57647 & 42353 & 52403 & 47597 & 14324\tabularnewline
56 & 60630 & 39370 & 55293 & 44707 & 14669\tabularnewline
57 & 63647 & 36353 & 57787 & 42213 & 15103\tabularnewline
58 & 66961 & 33039 & 60439 & 39561 & 15450\tabularnewline
59 & 69968 & 30032 & 63708 & 36292 & 15246\tabularnewline
60 & 72838 & 27162 & 66270 & 33730 & 15177\tabularnewline
61 & 75609 & 24391 & 68873 & 31127 & 15006\tabularnewline
62 & 77999 & 22001 & 71318 & 28682 & 14538\tabularnewline
63 & 80616 & 19384 & 73517 & 26483 & 14510\tabularnewline
64 & 83089 & 16911 & 76209 & 23791 & 13784\tabularnewline
65 & 85150 & 14850 & 78086 & 21914 & 13412\tabularnewline
66 & 86908 & 13092 & 80437 & 19563 & 12477\tabularnewline
67 & 88671 & 11329 & 82379 & 17621 & 11718\tabularnewline
68 & 90537 & 9463 & 84483 & 15517 & 10878\tabularnewline
69 & 91969 & 8031 & 86233 & 13767 & 10033\tabularnewline
70 & 93181 & 6819 & 87914 & 12086 & 9070\tabularnewline
71 & 94330 & 5670 & 89200 & 10800 & 8401\tabularnewline
72 & 95324 & 4676 & 90833 & 9167 & 7359\tabularnewline
73 & 96143 & 3857 & 91805 & 8195 & 6726\tabularnewline
74 & 96959 & 3041 & 93065 & 6935 & 5900\tabularnewline
75 & 97543 & 2457 & 93987 & 6013 & 5222\tabularnewline
76 & 97969 & 2031 & 94900 & 5100 & 4481\tabularnewline
77 & 98497 & 1503 & 95756 & 4244 & 3809\tabularnewline
78 & 98744 & 1256 & 96442 & 3558 & 3269\tabularnewline
79 & 99045 & 955 & 96942 & 3058 & 2837\tabularnewline
80 & 99276 & 724 & 97528 & 2472 & 2329\tabularnewline
81 & 99481 & 519 & 97996 & 2004 & 1894\tabularnewline
82 & 99556 & 444 & 98321 & 1679 & 1597\tabularnewline
83 & 99691 & 309 & 98722 & 1278 & 1227\tabularnewline
84 & 99752 & 248 & 98943 & 1057 & 1015\tabularnewline
85 & 99833 & 167 & 99144 & 856 & 837\tabularnewline
86 & 99895 & 105 & 99346 & 654 & 642\tabularnewline
87 & 99925 & 75 & 99461 & 539 & 530\tabularnewline
88 & 99945 & 55 & 99566 & 434 & 428\tabularnewline
89 & 99976 & 24 & 99675 & 325 & 324\tabularnewline
90 & 99977 & 23 & 99756 & 244 & 243\tabularnewline
91 & 99982 & 18 & 99839 & 161 & 155\tabularnewline
92 & 99988 & 12 & 99865 & 135 & 135\tabularnewline
93 & 99994 & 6 & 99885 & 115 & 115\tabularnewline
94 & 99993 & 7 & 99911 & 89 & 88\tabularnewline
95 & 99998 & 2 & 99953 & 47 & 47\tabularnewline
96 & 99999 & 1 & 99965 & 35 & 35\tabularnewline
97 & 99999 & 1 & 99979 & 21 & 21\tabularnewline
98 & 100000 & 0 & 99973 & 27 & 27\tabularnewline
99 & 100000 & 0 & 99984 & 16 & 16\tabularnewline
100 & 100000 & 0 & 99989 & 11 & 11\tabularnewline
\bottomrule
\end{longtable}

Overall, as
expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}},
predator-prey communities are relatively stable while mutualist
communties are highly unstable. But interestingly, while \(Var(\gamma)\)
stabilises predator-prey and competitor communities, it does not
stabilise mutualist communities. This is unsurprising because purely
mutualist communities are characterised by a very
positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
alone will shift all real parts of eigenvalues to negative values.

\hypertarget{connectance}{\section{Different inter-connectivity (C)
values}\label{connectance}}

In the main text, for simplicity, I assumed inter-connectivity values of
\(C = 1\), meaning that all off-diagonal elements of a matrix \(M\) were
potentially nonzero and sampled from a normal distribution
\(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here I present
four tables showing the number of stable communities given
\(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
component response time (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
higher number of stable communities than when \(\gamma\) did not vary
(\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
million \(M\) were simulated. As with the results on
\protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
in the tables below \texttt{A0} refers to matrices when \(\gamma = 1\),
and \texttt{A1} refers to matrices after \(Var(\gamma)\) is added. The
column \texttt{A0\_unstable} shows the number of \texttt{A0} matrices
that are unstable, and the column \texttt{A0\_stable} shows the number
of \texttt{A0} matrices that are stable (these two columns sum to
100000). Similarly, the column \texttt{A1\_unstable} shows the number of
\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
number that are stable. The columns \texttt{A1\_stabilised} and
\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
stabilised or destabilised, respectively, by \(Var(\gamma)\).

All data reported below for various values of \(C\) are accessible using
the below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{C3dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt3.csv"}\NormalTok{);}
\NormalTok{C5dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt5.csv"}\NormalTok{);}
\NormalTok{C7dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt7.csv"}\NormalTok{);}
\NormalTok{C9dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt9.csv"}\NormalTok{);}
\end{Highlighting}
\end{Shaded}

These objects \texttt{C3dat}, \texttt{C5dat}, \texttt{C7dat}, and
\texttt{C9dat} include the results for \(C = 0.3\), \(C = 0.5\),
\(C = 0.7\), and \(C = 0.9\), respectively.

\textbf{Connectance \(C = 0.3\)}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
3 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
4 & 24 & 99976 & 24 & 99976 & 0 & 0\tabularnewline
5 & 59 & 99941 & 59 & 99941 & 0 & 0\tabularnewline
6 & 98 & 99902 & 98 & 99902 & 0 & 0\tabularnewline
7 & 160 & 99840 & 161 & 99839 & 0 & 1\tabularnewline
8 & 290 & 99710 & 293 & 99707 & 0 & 3\tabularnewline
9 & 430 & 99570 & 434 & 99566 & 0 & 4\tabularnewline
10 & 648 & 99352 & 653 & 99347 & 1 & 6\tabularnewline
11 & 946 & 99054 & 957 & 99043 & 0 & 11\tabularnewline
12 & 1392 & 98608 & 1415 & 98585 & 4 & 27\tabularnewline
13 & 2032 & 97968 & 2065 & 97935 & 5 & 38\tabularnewline
14 & 2627 & 97373 & 2688 & 97312 & 10 & 71\tabularnewline
15 & 3588 & 96412 & 3647 & 96353 & 35 & 94\tabularnewline
16 & 5019 & 94981 & 5124 & 94876 & 51 & 156\tabularnewline
17 & 6512 & 93488 & 6673 & 93327 & 79 & 240\tabularnewline
18 & 8444 & 91556 & 8600 & 91400 & 165 & 321\tabularnewline
19 & 10416 & 89584 & 10667 & 89333 & 244 & 495\tabularnewline
20 & 13254 & 86746 & 13477 & 86523 & 425 & 648\tabularnewline
21 & 16248 & 83752 & 16481 & 83519 & 642 & 875\tabularnewline
22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151\tabularnewline
23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490\tabularnewline
24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818\tabularnewline
25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137\tabularnewline
26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506\tabularnewline
27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811\tabularnewline
28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072\tabularnewline
29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321\tabularnewline
30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480\tabularnewline
31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568\tabularnewline
32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531\tabularnewline
33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360\tabularnewline
34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014\tabularnewline
35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773\tabularnewline
36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393\tabularnewline
37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913\tabularnewline
38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566\tabularnewline
39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187\tabularnewline
40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936\tabularnewline
41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683\tabularnewline
42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449\tabularnewline
43 & 99275 & 725 & 98312 & 1688 & 1251 & 288\tabularnewline
44 & 99583 & 417 & 98872 & 1128 & 903 & 192\tabularnewline
45 & 99776 & 224 & 99339 & 661 & 576 & 139\tabularnewline
46 & 99865 & 135 & 99518 & 482 & 413 & 66\tabularnewline
47 & 99938 & 62 & 99744 & 256 & 226 & 32\tabularnewline
48 & 99956 & 44 & 99824 & 176 & 151 & 19\tabularnewline
49 & 99980 & 20 & 99914 & 86 & 85 & 19\tabularnewline
50 & 99993 & 7 & 99950 & 50 & 46 & 3\tabularnewline
51 & 99998 & 2 & 99971 & 29 & 28 & 1\tabularnewline
52 & 99998 & 2 & 99986 & 14 & 14 & 2\tabularnewline
53 & 99999 & 1 & 99992 & 8 & 7 & 0\tabularnewline
54 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
55 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
56 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
57 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
58 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
100 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Connectance \(C = 0.5\)}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
3 & 32 & 99968 & 32 & 99968 & 0 & 0\tabularnewline
4 & 122 & 99878 & 122 & 99878 & 0 & 0\tabularnewline
5 & 320 & 99680 & 321 & 99679 & 0 & 1\tabularnewline
6 & 667 & 99333 & 673 & 99327 & 0 & 6\tabularnewline
7 & 1233 & 98767 & 1252 & 98748 & 0 & 19\tabularnewline
8 & 2123 & 97877 & 2156 & 97844 & 3 & 36\tabularnewline
9 & 3415 & 96585 & 3471 & 96529 & 16 & 72\tabularnewline
10 & 5349 & 94651 & 5450 & 94550 & 30 & 131\tabularnewline
11 & 7990 & 92010 & 8185 & 91815 & 81 & 276\tabularnewline
12 & 11073 & 88927 & 11301 & 88699 & 219 & 447\tabularnewline
13 & 14971 & 85029 & 15204 & 84796 & 445 & 678\tabularnewline
14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002\tabularnewline
15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404\tabularnewline
16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980\tabularnewline
17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476\tabularnewline
18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937\tabularnewline
19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262\tabularnewline
20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527\tabularnewline
21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673\tabularnewline
22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491\tabularnewline
23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244\tabularnewline
24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876\tabularnewline
25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382\tabularnewline
26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901\tabularnewline
27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358\tabularnewline
28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974\tabularnewline
29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657\tabularnewline
30 & 99103 & 897 & 98033 & 1967 & 1466 & 396\tabularnewline
31 & 99502 & 498 & 98665 & 1335 & 1068 & 231\tabularnewline
32 & 99745 & 255 & 99185 & 815 & 696 & 136\tabularnewline
33 & 99881 & 119 & 99572 & 428 & 375 & 66\tabularnewline
34 & 99955 & 45 & 99788 & 212 & 191 & 24\tabularnewline
35 & 99979 & 21 & 99900 & 100 & 95 & 16\tabularnewline
36 & 99995 & 5 & 99950 & 50 & 50 & 5\tabularnewline
37 & 99997 & 3 & 99970 & 30 & 28 & 1\tabularnewline
38 & 99998 & 2 & 99986 & 14 & 13 & 1\tabularnewline
39 & 99999 & 1 & 99991 & 9 & 9 & 1\tabularnewline
40 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
41 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
42 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
43 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Connectance \(C = 0.7\)}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
3 & 106 & 99894 & 106 & 99894 & 0 & 0\tabularnewline
4 & 395 & 99605 & 397 & 99603 & 0 & 2\tabularnewline
5 & 1117 & 98883 & 1123 & 98877 & 0 & 6\tabularnewline
6 & 2346 & 97654 & 2367 & 97633 & 6 & 27\tabularnewline
7 & 4314 & 95686 & 4388 & 95612 & 16 & 90\tabularnewline
8 & 7327 & 92673 & 7456 & 92544 & 61 & 190\tabularnewline
9 & 11514 & 88486 & 11792 & 88208 & 150 & 428\tabularnewline
10 & 16247 & 83753 & 16584 & 83416 & 415 & 752\tabularnewline
11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162\tabularnewline
12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818\tabularnewline
13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355\tabularnewline
14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927\tabularnewline
15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421\tabularnewline
16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527\tabularnewline
17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495\tabularnewline
18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351\tabularnewline
19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979\tabularnewline
20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339\tabularnewline
21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869\tabularnewline
22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267\tabularnewline
23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791\tabularnewline
24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504\tabularnewline
25 & 99433 & 567 & 98630 & 1370 & 1079 & 276\tabularnewline
26 & 99760 & 240 & 99259 & 741 & 655 & 154\tabularnewline
27 & 99895 & 105 & 99576 & 424 & 377 & 58\tabularnewline
28 & 99950 & 50 & 99790 & 210 & 194 & 34\tabularnewline
29 & 99981 & 19 & 99915 & 85 & 80 & 14\tabularnewline
30 & 99994 & 6 & 99952 & 48 & 47 & 5\tabularnewline
31 & 99998 & 2 & 99972 & 28 & 28 & 2\tabularnewline
32 & 99999 & 1 & 99992 & 8 & 8 & 1\tabularnewline
33 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
34 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
35 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Connectance \(C = 0.9\)}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 14 & 99986 & 14 & 99986 & 0 & 0\tabularnewline
3 & 240 & 99760 & 240 & 99760 & 0 & 0\tabularnewline
4 & 1008 & 98992 & 1016 & 98984 & 0 & 8\tabularnewline
5 & 2708 & 97292 & 2729 & 97271 & 2 & 23\tabularnewline
6 & 5669 & 94331 & 5755 & 94245 & 13 & 99\tabularnewline
7 & 9848 & 90152 & 10057 & 89943 & 91 & 300\tabularnewline
8 & 15903 & 84097 & 16201 & 83799 & 336 & 634\tabularnewline
9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168\tabularnewline
10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852\tabularnewline
11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507\tabularnewline
12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127\tabularnewline
13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649\tabularnewline
14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765\tabularnewline
15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550\tabularnewline
16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092\tabularnewline
17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424\tabularnewline
18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746\tabularnewline
19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167\tabularnewline
20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729\tabularnewline
21 & 99111 & 889 & 98133 & 1867 & 1389 & 411\tabularnewline
22 & 99588 & 412 & 98905 & 1095 & 903 & 220\tabularnewline
23 & 99837 & 163 & 99480 & 520 & 452 & 95\tabularnewline
24 & 99932 & 68 & 99744 & 256 & 228 & 40\tabularnewline
25 & 99976 & 24 & 99863 & 137 & 133 & 20\tabularnewline
26 & 99995 & 5 & 99950 & 50 & 49 & 4\tabularnewline
27 & 99996 & 4 & 99986 & 14 & 13 & 3\tabularnewline
28 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
29 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
30 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
31 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{gam_dist}{\section{\texorpdfstring{Different distributions
of
\(\gamma\)}{Different distributions of \textbackslash{}gamma}}\label{gam_dist}}

In the main text, I considered a uniform distribution of component
response rates \(\gamma \sim \mathcal{U}(0, 2)\). The number of unstable
and stable \(M\) matrices are reported in \protect\hyperlink{IncrS}{a
table above} across different values of \(S\). Here I show complementary
results for three different distributions including an exponential,
beta, and gamma distribution of \(\gamma\) values. The shape of these
distributions is shown in the figure below.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Distributions of component response rate
(\(\boldsymbol{\gamma}\)) values in complex systems.} The stabilities of
simulated complex systems with these \(\gamma\) distributions are
compared to otherwise identical complex systems with a fixed component
response rate of \(\gamma = 1\) across different system sizes (\(S\);
i.e., component numbers) given a unit \(\gamma\) standard deviation
(\(\sigma_{\gamma} = 1\)) for b-d. Distributions are as follows: (a)
uniform, (b) exponential, (c) beta (\(\alpha = 0.5\) and
\(\beta = 0.5\)), and (d) gamma (\(k = 2\) and \(\theta = 2\)). Each
panel shows 1 million randomly generated \(\gamma\) values.

\includegraphics{SI_files/figure-latex/unnamed-chunk-23-1.pdf}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

The same 100000 \(M\) matrices were used to investigate stability when
applying each of these different distributions of \(\gamma\) values. The
table below shows the number of \(M\) that were unstable (\_unst) and
stable (\_stbl) for the exponential (Exp), beta, and gamma
distributions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fourdists <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/different_distr/four_distr_rand.csv"}\NormalTok{);}
\KeywordTok{kable}\NormalTok{(fourdists);}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
S & Exp\_unst & Exp\_stbl & beta\_unst & beta\_stbl & gamma\_unst &
gamma\_stbl\tabularnewline
\midrule
\endhead
2 & 30 & 99970 & 30 & 99970 & 30 & 99970\tabularnewline
3 & 355 & 99645 & 355 & 99645 & 355 & 99645\tabularnewline
4 & 1506 & 98494 & 1512 & 98488 & 1516 & 98484\tabularnewline
5 & 3930 & 96070 & 3971 & 96029 & 4006 & 95994\tabularnewline
6 & 7738 & 92262 & 7844 & 92156 & 7918 & 92082\tabularnewline
7 & 13606 & 86394 & 13889 & 86111 & 13990 & 86010\tabularnewline
8 & 20535 & 79465 & 21002 & 78998 & 21114 & 78886\tabularnewline
9 & 28614 & 71386 & 29060 & 70940 & 29110 & 70890\tabularnewline
10 & 38375 & 61625 & 38388 & 61612 & 38441 & 61559\tabularnewline
11 & 48616 & 51384 & 48211 & 51789 & 47957 & 52043\tabularnewline
12 & 59254 & 40746 & 58025 & 41975 & 57473 & 42527\tabularnewline
13 & 68816 & 31184 & 66753 & 33247 & 66127 & 33873\tabularnewline
14 & 77721 & 22279 & 75149 & 24851 & 74222 & 25778\tabularnewline
15 & 84842 & 15158 & 82030 & 17970 & 81040 & 18960\tabularnewline
16 & 90365 & 9635 & 87809 & 12191 & 86600 & 13400\tabularnewline
17 & 94171 & 5829 & 91756 & 8244 & 90668 & 9332\tabularnewline
18 & 96978 & 3022 & 94977 & 5023 & 94176 & 5824\tabularnewline
19 & 98376 & 1624 & 97018 & 2982 & 96268 & 3732\tabularnewline
20 & 99218 & 782 & 98357 & 1643 & 97765 & 2235\tabularnewline
21 & 99678 & 322 & 99124 & 876 & 98746 & 1254\tabularnewline
22 & 99864 & 136 & 99599 & 401 & 99323 & 677\tabularnewline
23 & 99954 & 46 & 99783 & 217 & 99668 & 332\tabularnewline
24 & 99978 & 22 & 99920 & 80 & 99821 & 179\tabularnewline
25 & 99996 & 4 & 99967 & 33 & 99911 & 89\tabularnewline
26 & 99999 & 1 & 99979 & 21 & 99960 & 40\tabularnewline
27 & 99999 & 1 & 99990 & 10 & 99983 & 17\tabularnewline
28 & 100000 & 0 & 99999 & 1 & 99991 & 9\tabularnewline
29 & 100000 & 0 & 99999 & 1 & 99999 & 1\tabularnewline
30 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
31 & 100000 & 0 & 100000 & 0 & 99999 & 1\tabularnewline
32 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
\bottomrule
\end{longtable}

In comparison to the uniform distribution (a), proportionally fewer
random systems are found with the exponential distribution (b), while
more are found with the beta (c) and gamma (d) distributions.

\hypertarget{ga}{\section{Genetic algorithm}\label{ga}}

Ideally, to investigate the potential of \(Var(\gamma)\) for increasing
the proportion of stable complex systems, the search space of all
possible \(\gamma\) vectors would be evaluated for each unique \(M\).
This is technically impossible because \(\gamma_{i}\) can take any real
value between 0-2, but even rounding \(\gamma\) to reasonable values
would result in a search space too large to practically explore. Under
these conditions, genetic algorithms are highly useful tools for finding
practical solutions by mimicking the process of biological
evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{6}}. In
this case, the practical solution is finding vectors of \(\gamma\) that
decrease the most positive real eigenvalue of \(M\). The genetic
algorithm below achieves this by initialising a large population of 1000
different potential \(\gamma\) vectors and allowing this population to
evolve through a process of mutation, crossover (swaping \(\gamma_{i}\)
values between vectors), selection, and reproduction until either a
\(\gamma\) vector is found where all \(\Re(\lambda) < 0\) or some
``giving up'' critiera is met (in the below, this ``giving up''"
criteria is met when 20 generations pass, or if the fitness increase
from one generation to the next is below a certain criteria). The
genetic algorithm relies on five functions. The first outer function
\texttt{Evo\_rand\_gen\_var} runs all of the simulations
(\texttt{max\_sp} refers to the maximum \(S\) value simulated, and
\texttt{iters} refers to the number of \(M\) to try for each \(S\)).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Evo_rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
        \NormalTok{nn             <-}\StringTok{ }\NormalTok{i;}
        \NormalTok{A1_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{A2_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{A1_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{A2_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{);}
        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{rand_mat_ga}\NormalTok{(A1);}
            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(i);}
    \NormalTok{\}}
    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
    \KeywordTok{return}\NormalTok{(all_res);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Note that \texttt{Evo\_rand\_gen\_var} calls three custom sub-functions,
\texttt{species\_interactions}, \texttt{rand\_mat\_ga}, and
\texttt{summarise\_randmat}. The first simply allows for non-random
interactions between components (e.g., modelling
\protect\hyperlink{ecological}{ecological interactions} of random,
competition, mutualism, or predator-prey).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
                \NormalTok{\}}
            \NormalTok{\}}
        \NormalTok{\}}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(mat);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The sub-function \texttt{rand\_mat\_ga} does the work of the genetic
algorithm, searching for \(\gamma\) vectors that are stabilising.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rand_mat_ga <-}\StringTok{ }\NormalTok{function(A1, }\DataTypeTok{max_it =} \DecValTok{20}\NormalTok{, }\DataTypeTok{converg =} \FloatTok{0.01}\NormalTok{)\{}
    \NormalTok{nn       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{];}
    \NormalTok{rind     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{1}\NormalTok{);}
    \NormalTok{inds     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{rind, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
    \NormalTok{lastf    <-}\StringTok{ }\NormalTok{-}\DecValTok{10}\NormalTok{;}
    \NormalTok{ccrit    <-}\StringTok{ }\DecValTok{10}\NormalTok{;}
    \NormalTok{find_st  <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
    \NormalTok{iter     <-}\StringTok{ }\NormalTok{max_it;}
    \NormalTok{while(iter >}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\NormalTok{find_st <}\StringTok{ }\DecValTok{1} \NormalTok{&}\StringTok{ }\NormalTok{ccrit >}\StringTok{ }\NormalTok{converg)\{}
        \NormalTok{ivar  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
        \NormalTok{ifit  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
        \NormalTok{isst  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{])\{}
            \NormalTok{ifit[i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*}\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values));}
            \NormalTok{ivar[i] <-}\StringTok{ }\KeywordTok{var}\NormalTok{(inds[i,]);}
            \NormalTok{isst[i] <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{\}}
        \NormalTok{most_fit <-}\StringTok{ }\KeywordTok{order}\NormalTok{(ifit, }\DataTypeTok{decreasing =} \OtherTok{TRUE}\NormalTok{)[}\DecValTok{1}\NormalTok{:}\DecValTok{20}\NormalTok{];}
        \NormalTok{parents  <-}\StringTok{ }\NormalTok{inds[most_fit,];}
        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{t}\NormalTok{(parents), }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn, }
                           \DataTypeTok{byrow =} \OtherTok{TRUE}\NormalTok{);}
        \NormalTok{mu_dat   <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.2}\NormalTok{);}
        \NormalTok{mu_dat2  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.02}\NormalTok{);}
        \NormalTok{mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{];}
        \NormalTok{mu_dat2[mu_dat2 >}\StringTok{ }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{2}\NormalTok{;}
        \NormalTok{mu_dat3  <-}\StringTok{ }\NormalTok{mu_dat *}\StringTok{ }\NormalTok{mu_dat2;}
        \NormalTok{mu_mat   <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{mu_dat3, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
        \NormalTok{new_gen  <-}\StringTok{ }\NormalTok{new_gen +}\StringTok{ }\NormalTok{mu_mat;}
        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{crossover}\NormalTok{(}\DataTypeTok{inds =} \NormalTok{new_gen, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{);}
        \NormalTok{inds     <-}\StringTok{ }\NormalTok{new_gen;}
        \NormalTok{find_st  <-}\StringTok{ }\KeywordTok{max}\NormalTok{(isst);}
        \NormalTok{newf     <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(ifit);}
        \NormalTok{ccrit    <-}\StringTok{ }\NormalTok{newf -}\StringTok{ }\NormalTok{lastf;}
        \NormalTok{lastf    <-}\StringTok{ }\NormalTok{newf;}
        \NormalTok{iter     <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{\}}
    \NormalTok{if(find_st ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
        \NormalTok{s_row <-}\StringTok{ }\KeywordTok{which}\NormalTok{(isst ==}\StringTok{ }\DecValTok{1}\NormalTok{)[}\DecValTok{1}\NormalTok{];}
        \NormalTok{writt <-}\StringTok{ }\KeywordTok{c}\NormalTok{(nn, inds[s_row,]);}
        \KeywordTok{cat}\NormalTok{(writt, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
        \KeywordTok{cat}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(find_st);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The while loop in \texttt{rand\_mat\_ga} continues until either
\texttt{iter} generations have occured, a solution \(\gamma\) vector is
found that results in all \(\Re(\lambda) < 0\), or some criteria of
minimum fitness increase is observed (by default,
\texttt{converg\ =\ 0.01}). Within the genetic algorithm, \(\gamma\)
values are mutated, crossover occurs between \(\gamma\) vectors, and
selection occurs in each generation such that the 20 \(\gamma\) vectors
that produce the lowest maximum \(\Re(\lambda)\) are allowed to have 50
offspring each. In mutation, any \(\gamma_{i}\) values that mutate below
zero are multiplied by \(-1\), and any values that mutate above 2 are
set to 2. Note also that if a solution is found, then one such
\(\gamma\) vector causing stability is printed to a file.

Crossover occurs in the \texttt{crossover} function below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{crossover <-}\StringTok{ }\NormalTok{function(inds, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{)\{}
    \NormalTok{crossed <-}\StringTok{ }\KeywordTok{floor}\NormalTok{(}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{] *}\StringTok{ }\NormalTok{pr);}
    \NormalTok{cross1  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
    \NormalTok{cross2  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(cross1))\{}
        \NormalTok{fromv   <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
        \NormalTok{tov     <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
        \NormalTok{temp                   <-}\StringTok{ }\NormalTok{inds[cross1[i],fromv:tov];}
        \NormalTok{inds[cross1[i],fromv:tov] <-}\StringTok{ }\NormalTok{inds[cross2[i],fromv:tov];}
        \NormalTok{inds[cross2[i],fromv:tov] <-}\StringTok{ }\NormalTok{temp;}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(inds);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

After all \(M\) are simulated in \texttt{Evo\_rand\_gen\_var}, the
\texttt{summarise\_randmat} formats the data into a table.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{summarise_randmat_ga <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{10}\NormalTok{);}
    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
        \NormalTok{unstables <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
        \NormalTok{stables   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{unstabled <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
        \NormalTok{stabled   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{non_feas  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
        \NormalTok{feasibl   <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{unfeased  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
        \NormalTok{feased    <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{foundd    <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{3}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstables);}
        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stables);}
        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstabled);}
        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stabled);}
        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(non_feas);}
        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feasibl);}
        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unfeased);}
        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feased);}
        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(foundd);}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(all_res);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Some stability results from this table are shown below. Each histogram
shows a different distribution of \(\gamma\) that was found to be
stabilising.

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 4 & 9996 & 4 & 9996 & 0 & 0\tabularnewline
3 & 42 & 9958 & 42 & 9958 & 0 & 0\tabularnewline
4 & 133 & 9867 & 133 & 9867 & 0 & 0\tabularnewline
5 & 414 & 9586 & 411 & 9589 & 3 & 0\tabularnewline
6 & 809 & 9191 & 799 & 9201 & 10 & 0\tabularnewline
7 & 1380 & 8620 & 1339 & 8661 & 41 & 0\tabularnewline
8 & 2074 & 7926 & 1927 & 8073 & 147 & 0\tabularnewline
9 & 2885 & 7115 & 2503 & 7497 & 382 & 0\tabularnewline
10 & 3842 & 6158 & 3158 & 6842 & 684 & 0\tabularnewline
11 & 4867 & 5133 & 3613 & 6387 & 1255 & 1\tabularnewline
12 & 5932 & 4068 & 4148 & 5852 & 1784 & 0\tabularnewline
13 & 6937 & 3063 & 4470 & 5530 & 2468 & 1\tabularnewline
14 & 7784 & 2216 & 4724 & 5276 & 3060 & 0\tabularnewline
15 & 8519 & 1481 & 5086 & 4914 & 3433 & 0\tabularnewline
16 & 9081 & 919 & 5262 & 4738 & 3819 & 0\tabularnewline
17 & 9431 & 569 & 5368 & 4632 & 4063 & 0\tabularnewline
18 & 9671 & 329 & 5571 & 4429 & 4100 & 0\tabularnewline
19 & 9844 & 156 & 5807 & 4193 & 4037 & 0\tabularnewline
20 & 9934 & 66 & 6133 & 3867 & 3801 & 0\tabularnewline
21 & 6387 & 34 & 6421 & 3579 & 3545 & 0\tabularnewline
22 & 6634 & 11 & 6645 & 3355 & 3344 & 0\tabularnewline
23 & 7037 & 8 & 7045 & 2955 & 2947 & 0\tabularnewline
24 & 7468 & 3 & 7471 & 2529 & 2526 & 0\tabularnewline
25 & 7816 & 0 & 7816 & 2184 & 2184 & 0\tabularnewline
26 & 8192 & 0 & 8192 & 1808 & 1808 & 0\tabularnewline
27 & 8680 & 0 & 8680 & 1320 & 1320 & 0\tabularnewline
28 & 8936 & 0 & 8936 & 1064 & 1064 & 0\tabularnewline
29 & 9296 & 0 & 9296 & 704 & 704 & 0\tabularnewline
30 & 9523 & 0 & 9523 & 477 & 477 & 0\tabularnewline
31 & 9705 & 0 & 9705 & 295 & 295 & 0\tabularnewline
32 & 9816 & 0 & 9816 & 184 & 184 & 0\tabularnewline
33 & 9894 & 0 & 9894 & 106 & 106 & 0\tabularnewline
34 & 9941 & 0 & 9941 & 59 & 59 & 0\tabularnewline
35 & 9968 & 0 & 9968 & 32 & 32 & 0\tabularnewline
36 & 9991 & 0 & 9991 & 9 & 9 & 0\tabularnewline
37 & 9993 & 0 & 9993 & 7 & 7 & 0\tabularnewline
38 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
39 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
40 & 10000 & 0 & 10000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

The distributions of nine \(\gamma\) vectors from the highest \(S\)
values are shown below. Recall that 1 million random matrices were
generated for the less computationally intense task of
\protect\hyperlink{IncrS}{comparing} \(M\) when \(\gamma = 1\) versus
when \(\gamma \sim \mathcal{U}(0, 2)\), so it is more informative to
compare stability in column 5 above with column 3 above. \textbf{This
comparison shows the high number of stable \(M\) that can be produced
through a targetted search of \(\gamma\) values, and suggests that many
otherwise unstable systems could potentially be stabilised by an
informed manipulation of their component response times. Such a
possibility might conceivably reduce the dimensionality of problems
involving stability in social-ecological or economic systems.}

Distributions of \(\gamma\) values in vectors for the highest values of
\(S\) are shown below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{evo_out <-}\StringTok{ }\KeywordTok{scan}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/evolved/evo_out.txt"}\NormalTok{);}
\KeywordTok{plot_evo_out}\NormalTok{(evo_out);}
\end{Highlighting}
\end{Shaded}

\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}

The distribution of \(\gamma\) values found by the genetic algorithm is
uniform. A uniform distribution was used to initialise \(\gamma\)
values, so there is therefore no evidence that a particular distribution
of \(\gamma\) is likely to be found to stabilise a matrix \(M\).

\hypertarget{Feasibility}{\section{Feasibility of complex
systems}\label{Feasibility}}

For complex systems in which individual system components (\(S\))
represent the density of some tangible quantity, it is important to
consider the feasibility of the system. Feasibile equilibria assume that
the values of all system components are positive at
equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{7}--\protect\hyperlink{ref-Song2018}{9}}.
This is of particular interest for ecological communities because
population density cannot take negative values, meaning that ecological
systems need to be feasible for stability to be biologically
realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
Consequently, the use of random matrices and traditional stability
critiera for making inferences in theoretical analyses of species
networks has recently been
criticised\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
While the key results in the main text are intended to be general to all
complex systems, and not restricted to species networks, I have also
performed a feasibility analysis on all matrices \(M\). This analysis
reveals that feasibility is not affected by \(Var(\gamma)\), meaning
that for pure interacting species networks, variation in component
response time (i.e., species generation time) does not affect stability
at biologically realistic species densities. Nevertheless, ecological
interactions do not exist in isolation in empirical systems, but instead
interact with
evolutionary\textsuperscript{\protect\hyperlink{ref-Patel2018}{10}},
abiotic, or social-economic systems. The relevance of \(\gamma\) for
complex system stability presented in the main text should therefore not
be ignored in the broader context of ecological communities.

Dougoud et al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}
define the following feasibility criteria for ecological systems
characterised by \(S\) interacting species with varying densities.

\[x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.\]

In the above, \(x^{*}\) is the vector of species abundances at
equilibrium (for feasibility, all values in \(x^{*}\) must be positive).
The matrix \(I\) is the identity matrix (1s on the diagonal, 0s on the
off-diagonal elements), and the value \(\theta\) is strength of
intraspecific competition (diagonal elements). As I have done elsewhere,
diagonal values are set to \(-1\), so \(\theta = -1\). The variable
\(C\) is the inter-connectivity (i.e., `connectance') of the community,
which was set to \(C = 1\) throughout the manuscript and supplemental
information, except \protect\hyperlink{connectance}{where otherwise
noted}. The variable \(\delta\) is a normalisation parameter that
modulates the strength of interactions (\(\sigma\) in the main text),
which are held in \(A\). In the main text, implicitly, \(\delta = 0\)
underlying strong interactions. Hence, the whole \((CS)^{-\delta} = 1\),
so in the above, a diagonal matrix of -1s (\(\theta I\)) is added to
\(A\), which has a diagonal of all zeros and an off-diagonal affecting
species interactions (i.e., the expression \((CS)^{-\delta}\) relates to
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} stability
criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}} by
\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and hence
\((CS)^{-\delta} = 1\) for the randomly simulated systems in the main
text and supplemental information). The above criteria is therefore
reduced to the below; note that the parenthetical in both equations
produces an \(M\) matrix as used throughout the main text and
supplemental information,

\[x^{*} = -\left(\theta I + A\right)^{-1}r.\]

To check the feasibility criteria, I therefore inverted
\(M = (\theta I + A)\) and multiplied elements by -1, then multiplied
the resulting matrix by the vector of population growth rates \(r\).
Feasibility is satisfied if all of the elements of the resulting vector
are positive.

The population growth rate for an individual species \(i\) is sampled
from a normal distribution of \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\), as
shown in the \texttt{rand\_gen\_var} function in
\protect\hyperlink{IncrS}{the section} on ``Stability across increasing
\(S\)'' above. Hence, each component \(i\) of the complex system \(M\)
is assumed to be a species with a growth rate of \(r_{i}\). Note that
negative intrinsic growth rates are not unrealistic, and will occur in
obligate mutualists in the absence of a partner.

When feasibility was evaluated with and without variation in \(\gamma\),
there was no increase in stability for \(M\) where \(\gamma\) varied as
compared to where \(\gamma = 1\). Results below illustrate this result,
which was general to all other simulations performed.

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
S & A0\_infeasible & A0\_feasible & A1\_infeasible & A1\_feasible &
A1\_made\_feasible & A1\_made\_infeasible\tabularnewline
\midrule
\endhead
2 & 749978 & 250022 & 749942 & 250058 & 35552 & 35516\tabularnewline
3 & 874519 & 125481 & 874296 & 125704 & 36803 & 36580\tabularnewline
4 & 937192 & 62808 & 937215 & 62785 & 26440 & 26463\tabularnewline
5 & 968776 & 31224 & 968639 & 31361 & 16319 & 16182\tabularnewline
6 & 984313 & 15687 & 984463 & 15537 & 9006 & 9156\tabularnewline
7 & 992149 & 7851 & 992161 & 7839 & 4991 & 5003\tabularnewline
8 & 996124 & 3876 & 996103 & 3897 & 2644 & 2623\tabularnewline
9 & 998014 & 1986 & 998027 & 1973 & 1361 & 1374\tabularnewline
10 & 999031 & 969 & 999040 & 960 & 698 & 707\tabularnewline
11 & 999546 & 454 & 999514 & 486 & 377 & 345\tabularnewline
12 & 999764 & 236 & 999792 & 208 & 160 & 188\tabularnewline
13 & 999883 & 117 & 999865 & 135 & 105 & 87\tabularnewline
14 & 999938 & 62 & 999945 & 55 & 40 & 47\tabularnewline
15 & 999971 & 29 & 999964 & 36 & 31 & 24\tabularnewline
16 & 999988 & 12 & 999991 & 9 & 8 & 11\tabularnewline
17 & 999996 & 4 & 999991 & 9 & 8 & 3\tabularnewline
18 & 999997 & 3 & 999999 & 1 & 1 & 3\tabularnewline
19 & 999998 & 2 & 999997 & 3 & 3 & 2\tabularnewline
20 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
21 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
22 & 999999 & 1 & 1000000 & 0 & 0 & 1\tabularnewline
23 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
24 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
25 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
26 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
27 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
28 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
29 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
30 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
31 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
in pure species interaction networks.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-May1972}{}
1. May, R. M. Will a large complex system be stable? \emph{Nature}
\textbf{238,} 413--414 (1972).

\hypertarget{ref-Allesina2012}{}
2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
\emph{Nature} \textbf{483,} 205--208 (2012).

\hypertarget{ref-Allesina2015a}{}
3. Allesina, S. \& Tang, S. The stability--complexity relationship at
age 40: a random matrix perspective. \emph{Population Ecology} 63--75
(2015).
doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}

\hypertarget{ref-Tang2014b}{}
4. Tang, S. \& Allesina, S. Reactivity and stability of large
ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
(2014).

\hypertarget{ref-Allesina2011}{}
5. Allesina, S. \& Levine, J. M. A competitive network theory of species
diversity. \emph{Proceedings of the National Academy of Sciences of the
United States of America} \textbf{108,} 5638--5642 (2011).

\hypertarget{ref-Hamblin2013}{}
6. Hamblin, S. On the practical usage of genetic algorithms in ecology
and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
184--194 (2013).

\hypertarget{ref-Grilli2017}{}
7. Grilli, J. \emph{et al.} Feasibility and coexistence of large
ecological communities. \emph{Nature Communications} \textbf{8,} (2017).

\hypertarget{ref-Dougoud2018}{}
8. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
The feasibility of equilibria in large ecosystems: a primary but
neglected concept in the complexity-stability debate. \emph{PLOS
Computational Biology} \textbf{14,} e1005988 (2018).

\hypertarget{ref-Song2018}{}
9. Song, C. \& Saavedra, S. Will a small randomly assembled community be
feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).

\hypertarget{ref-Patel2018}{}
10. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
effects of eco-evolutionary feedbacks on community stability.
\emph{American Naturalist} \textbf{191,} 1--29 (2018).

\end{document}