### 2020-01-10 notebook/revision_notes.Rmd
Starting to write up the general idea

---
title: "Component response rate variation underlies the stability of complex systems (revision notes)"
author: "A. Bradley Duthie    ( alexander.duthie@stir.ac.uk )"
date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
  FK9 4LA
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - nature.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
biblio-style: apalike
---

Role of correlated matrices in stabilisation
================================================================================

```{r, echo = FALSE}
source("../R/random_mats.R");
source("../R/small_world.R");
rand_rho_var <- function(S, rhos, iters, int_type = 0, rmx = 0.4, C = 1, 
                         by = 1, sigma = 0.4){
    tot_res <- NULL;
    fea_res <- NULL;
    rho_res <- NULL;
    cmplxty <- NULL;
    for(i in 1:length(rhos)){
        iter           <- iters;
        tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 8);
        fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 6);
        cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
        while(iter > 0){
            r_vec    <- rnorm(n = S, mean = 0, sd = rmx);
            A0       <- build_rho_mat(S = S, sigma = sigma, rho = rhos[i]);
            gam1     <- runif(n = S, min = 0, max = 2);
            A1       <- A0 * gam1;
            A0       <- A0 * mean(gam1);
            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
            A0_rho   <- mat_rho(A0);
            A1_rho   <- mat_rho(A1);
            rho_diff <- A1_rho - A0_rho;
            rho_abs  <- abs(A1_rho) - abs(A0_rho);
            if(A0_stb == TRUE){
                tot_res[[i]][iter, 1] <- 1;
            }
            if(A1_stb == TRUE){
                tot_res[[i]][iter, 2] <- 1;
            }
            if(A0_fea == TRUE){
                fea_res[[i]][iter, 1] <- 1;
            }
            if(A1_fea == TRUE){
                fea_res[[i]][iter, 2] <- 1;
            }
            rho_res[[i]][iter, 1] <- A0_rho;
            rho_res[[i]][iter, 2] <- A1_rho;
            rho_res[[i]][iter, 3] <- rho_diff;
            rho_res[[i]][iter, 4] <- rho_abs;
            rho_res[[i]][iter, 5] <- max(Re(eigen(A0)$values));
            rho_res[[i]][iter, 6] <- max(Re(eigen(A1)$values));
            cmplxty[[i]][iter, 1] <- get_complexity(A0);
            cmplxty[[i]][iter, 2] <- get_complexity(A1);
            iter                  <- iter - 1;
        }
    }
    all_res     <- summarise_randmat(tot_res = tot_res, fea_res = fea_res,
                                     rho_res = rho_res, cmplxty = cmplxty);
    all_res[,1]          <- rhos;
    colnames(all_res)[1] <- "rho_set";
    return(all_res);
}

simpleboot <- function(freqs,repli=1000,alpha=0.05){    
    vals  <- NULL;                                                          
    i     <- 0;                                         
    while(i < repli){                                                   
        boot  <- sample(x=freqs,size=length(freqs),replace=TRUE);           
        strap <- mean(boot);                                                
        vals  <- c(vals,strap);                                             
        i     <- i + 1;                                                     
    }                                                                       
    vals   <- sort(x=vals,decreasing=FALSE);                            
    lowCI  <- vals[round((alpha*0.5)*repli)];                               
    highCI <- vals[round((1-(alpha*0.5))*repli)];                           
    CIs    <- c(lowCI,highCI);                                          
    return(CIs);                                                            
}

summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
    sims    <- length(tot_res);
    all_res <- matrix(data = 0, nrow = sims, ncol = 25);
    for(i in 1:sims){
        all_res[i, 1]  <- i + 1;
        # Stable and unstable
        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
        # Stabilised and destabilised
        all_res[i, 6] <- sum(tot_res[[i]][,1] == FALSE & 
                                 tot_res[[i]][,2] == TRUE);
        all_res[i, 7] <- sum(tot_res[[i]][,1] == TRUE & 
                                 tot_res[[i]][,2] == FALSE);
        # Feasible and infeasible
        all_res[i, 8]  <- sum(fea_res[[i]][,1] == FALSE);
        all_res[i, 9]  <- sum(fea_res[[i]][,1] == TRUE);
        all_res[i, 10]  <- sum(fea_res[[i]][,2] == FALSE);
        all_res[i, 11]  <- sum(fea_res[[i]][,2] == TRUE);
        # Feased and defeased
        all_res[i, 12] <- sum(fea_res[[i]][,1] == FALSE & 
                                  fea_res[[i]][,2] == TRUE);
        all_res[i, 13] <- sum(fea_res[[i]][,1] == TRUE & 
                                  fea_res[[i]][,2] == FALSE);
        # Mean correlations between A[i,j] and A[j,i] off-diag elements
        all_res[i, 14] <- mean(rho_res[[i]][,1]);
        all_res[i, 15] <- mean(rho_res[[i]][,2]);
        all_res[i, 16] <- mean(rho_res[[i]][,3]);
        all_res[i, 17] <- mean(rho_res[[i]][,4]);
        all_res[i, 18] <- mean(cmplxty[[i]][,1]);
        all_res[i, 19] <- mean(cmplxty[[i]][,2]);
        all_res[i, 20] <- mean(rho_res[[i]][,5]);
        all_res[i, 21] <- mean(rho_res[[i]][,6]);
        all_res[i, 22:23] <- range(rho_res[[i]][,5]);
        all_res[i, 24:25] <- range(rho_res[[i]][,6]);
    }
    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
                "A1_stabilised", "A1_destabilised", "A0_infeasible", 
                "A0_feasible", "A1_infeasible", "A1_feasible", 
                "A1_made_feasible", "A1_made_infeasible", "A0_rho", "A1_rho",
                "rho_diff", "rho_abs", "complex_A0", "complex_A1", "A0_eig",
                "A1_eig", "LCI_A0", "UCI_A0", "LCI_A1", "UCI_A1");
    colnames(all_res) <- cnames;
    return(all_res);
}
```

In complex systems represented by large random matrices, correlation between matrix elements $A_{ij}$ and $A_{ji}$ affects the distribution of eigenvalues and therefore local stability. As the correlation between matrix elements ($\rho$) decreases, the eigenvalue spectra changes such that more variation falls along the imaginary axis. The figure panels below compare a random matrix in which $\rho = 0$ (left) to one in which ($\rho$ = -0.5). In both cases, complex systems include $S = 1000$ components, with diagonal elements of $-1$ and off-diagonal elements drawn from a normal distribution with a mean of $0$ and $\sigma = 0.4$.

```{r, echo = FALSE}
mat_rho0  <- build_rho_mat(S = 1000, sigma = 0.4, rho = 0);
mat_rho5  <- build_rho_mat(S = 1000, sigma = 0.4, rho = -0.5);
mr0_eig   <- eigen(mat_rho0)$values;
mr5_eig   <- eigen(mat_rho5)$values;
mr0_r     <- Re(mr0_eig);
mr0_i     <- Im(mr0_eig);
mr5_r     <- Re(mr5_eig);
mr5_i     <- Im(mr5_eig);
par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0.2, 0.2));
plot(x = mr0_r, y = mr0_i, pch = 4, cex = 0.7, col = "firebrick",
     xlab = "", ylab = "", cex.lab = 1.25, cex.axis = 1.25, asp = 1, 
     ylim = c(min(mr5_i), max(mr5_i)));
plot(x = mr5_r, y = mr5_i, pch = 4, cex = 0.7, col = "firebrick", 
     xlab = "", ylab = "", cex.lab = 1.25, cex.axis = 1.25, asp = 1, 
     yaxt = "n");
mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 1.5);
mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 1.5);
```

Because of this affect of $\rho$ on the eigenvalue spectra, decreasing values of $\rho$ will also decrease the rightmost eigenvalue of the matrix $A$. This makes it more likely that the complex system represented by $A$ is locally stable, as stability occurs when all real parts of eigenvalues are negative. Note that this elongation along the imaginary axis is also characteristic of predator-prey communities (in which, by definition $A_{ij}$ and $A_{ji}$ have opposing signs). Also note that as $\rho$ increases such that $\rho > 0$, the same elongation happens along the real axis, making random complex systems less likely to be stable.

```{r, echo = FALSE}
sim_rhos <- rand_rho_var(S = 25, rhos = seq(from = -0.9, to = 0.9, by = 0.1), 
                         iters = 1000);
```

A simple numerical analysis illustrates the linear relationship between $\rho$ and the expected value of the real part of the leading eigenvalue $\max(\Re(\lambda))$. Below, I have run 1000 simulations across values of $\rho$ from `r min(sim_rhos[,1])` to `r max(sim_rhos[,1])` for complex systems with $S = 25$ components. Error bars show 95\% bootstrapped confidence intervals.

```{r, echo = FALSE}
par(mar = c(5, 5, 1, 1));
plot(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
     xlab = expression(paste("Correlation between A"[ij]," and A"[ji])),
     ylab = "Leading real part eigenvalue", 
     ylim = c(min(sim[,22]), max(sim[,23])));
arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
       y1 = sim_rhos[,22], angle = 90, length = 0.05, lwd = 2);
arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
       y1 = sim_rhos[,23], angle = 90, length = 0.05, lwd = 2);
```

```{r, echo = FALSE}
dat <- read.csv(file = "sim_results/C_1/random_all.csv");
dat <- dat[,-1];
```

In the main text, I demonstrated that when $S$ is finite but system complexity $\sigma\sqrt{SC}$ is high ($C$ defines the connectance of $A$, or the proportion of non-zero off-diagonal elements), variation in component response rate $\gamma$ often underlies system stability. Mathematically, this means that by multiplying $A$ by a diagonal matrix $\gamma$ with variable elements, the sign of $\max(\Re(\lambda))$ is sometimes flipped from positive to negative in these finite systems of high complexity. Interestingly, this increase in stability given $Var(\gamma) > 0$ is not caused by $\gamma$ decrease $\rho$. In fact, for $S = 25$, $\sigma = 0.4$, and $C = 1$, random complex systems that are stabilised by $\gamma$ typically have increased $\rho$ values. Note that for these parameter values, 1 million simulations found $M = \gamma A$ to be stable for `r dat[24, 3]` systems when all $\gamma_{i} = 1$, but `r dat[24, 5]` systems when $\gamma_{i} \sim \mathcal{U}(0, 2)$. Below shows the distribution of the difference in $\rho$ between systems with versus without $var(\gamma)$ for stabilised systems; that is, $\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}$.