### 2020-04-14 notebook/cover_letter.Rmd
Add changes for minor revision for Sci Rep

---
output:
  pdf_document: default
---

\fontsize{15}{22}
\selectfont

Dear Scientific Reports Editorial Board,

Please find attached my revision of the manuscript, "Component response rate variation underlies the stability of highly complex finite systems", which I hope you will further consider for publication as an Article.

My revision now addresses all remaining concerns from the previous round of review.

In the interest of transparency, the entire history of this project, including all my efforts in revising and addressing reviewer comments, is accessible on GitHub (https://github.com/bradduthie/RandomMatrixStability). An accompanying R package and website (https://bradduthie.github.io/RandomMatrixStability/) is also available with documentation for easily recreating all simulations and analyses.

I certify that this manuscript is original work and not under review at any other journal or book; a pre-print version of this revised manuscript has been submitted to arXiv (http://arxiv.org/abs/1806.01029).

Sincerely,

A. Bradley Duthie

University of Stirling, Stirling, UK