### 2020-04-14 notebook/sci_rep_reviewer_3.Rmd

---
title: "Reviewer Comments (Scientific Reports)"
author: "Brad Duthie"
output:
  pdf_document: default
  html_document: default
---

Reviewer comments
================================================================================

Reviewer #3 (Technical Comments to the Author):
--------------------------------------------------------------------------------

The author has followed the reviewers' recommendations and, in my opinion, has greatly improved the manuscript. I recommend accepting to be published.

Author response to Reviewer #3 comment
--------------------------------------------------------------------------------

I thank Reviewer 3 for their kind words, and for reviewing this manuscript a second time.

Reviewer #3 minor comment
--------------------------------------------------------------------------------

Minor comment

in lines 217 and 226 the article has lost el citation of Patel et al.

Author response to Reviewer #3 minor comment
--------------------------------------------------------------------------------

I have re-inserted the citation on these lines.